### 2022-02-22 paper.qmd
a first attempt

---
# Quarto-specific stuff you probably don't want to change for now
template: tvcg.latex
pdf-engine: pdflatex
documentclass: vgtc
format: pdf
cite-method: natbib
natbiboptions: square, numbers

# Stuff you'll want to change
title: Global Illumination for Fun and Profit
shortauthortitle: "Biv et al.: Global Illumination for Fun and Profit"
author:
  - Roy G. Biv
  - Ed Grimley, *Member, IEEE*
  - Martha Stewart
abstract: |
  Duis autem vel eum iriure dolor in hendrerit in vulputate
  velit esse molestie consequat, vel illum dolore eu feugiat nulla
  facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
  praesent luptatum zzril delenit augue duis dolore te feugait nulla
  facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
  sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
  aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
  vel eum iriure dolor in hendrerit in vulputate velit esse molestie
  consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
  accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
  delenit augue duis dolore te feugait nulla facilisi.
bibliography: paper.bib
firstsection: Introduction
tvcg-keywords:
  - Radiosity
  - global illumination
  - constant time
---

This template is for papers of VGTC-sponsored conferences such as IEEE
VIS, IEEE VR, and ISMAR which are published as special issues of
TVCG. The template does not contain the respective dates of the
conference/journal issue, these will be entered by IEEE as part of the
publication production process. Therefore, **please leave the
copyright statement at the bottom-left of this first page untouched**.

# Using the Sytle Template

* If you receive compilation errors along the lines of `Package ifpdf
  Error: Name clash, \ifpdf is already defined` then please add a new
  line `\let\ifpdf\relax` right after the
  `\documentclass[journal]{vgtc}` call. Note that your error is due to
  packages you use that define "`\ifpdf`" which is obsolete (the
  result is that `\ifpdf` is defined twice); these packages should be
  changed to use ifpdf package instead.

* Note that each author's affiliations have to be provided in the
  author footer on the bottom-left corner of the first page. It is
  permitted to merge two or more people from the same institution as
  long as they are shown in the same order as in the overall author
  sequence on the top of the first page. For example, if authors A, B,
  C, and D are from institutions 1, 2, 1, and 2, respectively, then it
  is ok to use 2 bullets as follows:

  * A and C are with Institution 1. E-mail: {a, c}@i1.com .
  * B and D are with Institution 2. E-mail: {b, d}@i2.org .

# Method

# Experiments

# Discussion

# Conclusions and Future Work