### 2022-02-22 paper.qmd
tables, equation numbers

---
# Quarto-specific stuff you probably don't want to change for now
pdf-engine: pdflatex
documentclass: vgtc
cite-method: natbib
natbiboptions: square, numbers
format:
  pdf:
    template: tvcg.latex
  latex:
    template: tvcg.latex
#  Uncomment the two lines below to also generate an HTML version
#  html:
#    dummy: unused

# Stuff you'll want to change
title: Global Illumination for Fun and Profit
shortauthortitle: "Biv et al.: Global Illumination for Fun and Profit"
author:
  - Roy G. Biv
  - Ed Grimley, *Member, IEEE*
  - Martha Stewart
abstract: |
  Duis autem vel eum iriure dolor in hendrerit in vulputate
  velit esse molestie consequat, vel illum dolore eu feugiat nulla
  facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
  praesent luptatum zzril delenit augue duis dolore te feugait nulla
  facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
  sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
  aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
  vel eum iriure dolor in hendrerit in vulputate velit esse molestie
  consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
  accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
  delenit augue duis dolore te feugait nulla facilisi.
bibliography: paper.bib
firstsection: Introduction
keyword:
  - Radiosity
  - global illumination
  - constant time
---

This template is for papers of VGTC-sponsored conferences such as IEEE
VIS, IEEE VR, and ISMAR which are published as special issues of
TVCG. The template does not contain the respective dates of the
conference/journal issue, these will be entered by IEEE as part of the
publication production process. Therefore, **please leave the
copyright statement at the bottom-left of this first page untouched**.

# Using the Style Template

(This information is incorrect for the [quarto](https://quarto.org)
template. We provide it below mostly as a markdown illustration.)

* If you receive compilation errors along the lines of `Package ifpdf
  Error: Name clash, \ifpdf is already defined` then please add a new
  line `\let\ifpdf\relax` right after the
  `\documentclass[journal]{vgtc}` call. Note that your error is due to
  packages you use that define "`\ifpdf`" which is obsolete (the
  result is that `\ifpdf` is defined twice); these packages should be
  changed to use ifpdf package instead.

* Note that each author's affiliations have to be provided in the
  author footer on the bottom-left corner of the first page. It is
  permitted to merge two or more people from the same institution as
  long as they are shown in the same order as in the overall author
  sequence on the top of the first page. For example, if authors A, B,
  C, and D are from institutions 1, 2, 1, and 2, respectively, then it
  is ok to use 2 bullets as follows:

  * A and C are with Institution 1. E-mail: {a, c}\@i1.com .
  * B and D are with Institution 2. E-mail: {b, d}\@i2.org .

* The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the "`\autoref{reference}`" call (instead of "`Figure\~{}\ref{reference}`" or similar) since "`\autoref{reference}`" turns the entire reference into an internal link, not just the number. Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.

* The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use "`\includegraphics{CypressView}`" (instead of "`\includegraphics{pictures/CypressView.jpg}`").

* For adding hyperlinks and DOIs to the list of references, you can use "`\bibliographystyle{abbrv-doi-hyperref-narrow}`" (instead of "`\bibliographystyle{abbrv}`"). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the "`http://dx.doi.org/`" url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]

  **Note 1**: occasionally (for some LaTeX distributions) this hyper-linked bibTeX style may lead to **compilation errors** ("`pdfendlink ended up in different nesting level ...`") if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to "`\bibliographystyle{abbrv-doi-narrow}`" (at the expense of removing hyperlinks from the bibliography) and try "`\bibliographystyle{abbrv-doi-hyperref-narrow}`" again after some more editing.

  **Note 2**: the "`-narrow`" versions of the bibliography style use the font "PTSansNarrow-TLF" for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the ["paratype" package](https://www.ctan.org/pkg/paratype), and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a "`-narrow`" bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the "`-narrow`" in the file name.

  DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the `abbrv-doi` bibliography style but the compiled PDF shows you the effect of the `abbrv-doi-hyperref-narrow` style.

# Bibliography Instructions

# Example Section

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam
voluptua @Kitware:2003,Max:1995:OMF. At vero eos et accusam et
justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est.

# Exposition

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat @Kindlmann:1999:SAG.

$$\sum_{j=1}^{z} j = \frac{z(z+1)}{2}$$ {#eq-1}

Lorem ipsum dolor sit amet (@eq-1), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

## Lorem Ipsum

Lorem ipsum dolor sit amet (see @tbl-vis-papers), consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum.

TBF: Finish this table.

| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: VIS/VisWeek accepted/presented papers from 1990--2016. {#tbl-vis-papers}

# Discussion

# Conclusions and Future Work