### 2022-02-22 paper.qmd
use CSL/citeproc instead of natbib

---
# Quarto-specific stuff you probably don't want to change for now
pdf-engine: pdflatex
csl: vgtc-tvcg.csl
link-citations: true

################################################################################

format:
  pdf:
    template: vgtc-quarto-template/tvcg.latex
#  Uncomment the line below to also generate an HTML version
#  html: default

################################################################################
# Paper metadata

title: Global Illumination for Fun and Profit
shortauthortitle: "Biv et al.: Global Illumination for Fun and Profit"
author:
  - name: Roy G. Biv
    institution: Starbucks Research
    email: "roy.g.biv@aol.com"
  - name: Ed Grimley
    institution: Grimley Widgets, Inc.
    email: "ed.grimley@aol.com"
  - name: Martha Stewart
    institution: Martha Stewart Enterprises at Microsoft Research
    email: "martha.stewart@marthastewart.com"

# If you are submitting a paper to a conference for review with a double
# blind reviewing process, please uncomment the line below and use your
# OnlineID. 
#online-id: 1337
# comment this line for final version
review: true

# please declare the paper type of your paper to help reviewers, only shown in review mode
# choices:
#   algorithm/technique
#   application/design study
#   evaluation
#   system
#   theory/model
paper-type: please specify

# Uncomment this to add a teaser to your paper

teaser:
  image-options: width=\linewidth
  image: CypressView.jpg
  caption: |
    In the Clouds: Vancouver from Cypress Mountain. Note that the
    teaser may not be wider than the abstract block.
  label: fig-teaser

abstract: |
  Duis autem vel eum iriure dolor in hendrerit in vulputate
  velit esse molestie consequat, vel illum dolore eu feugiat nulla
  facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
  praesent luptatum zzril delenit augue duis dolore te feugait nulla
  facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
  sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
  aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
  vel eum iriure dolor in hendrerit in vulputate velit esse molestie
  consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
  accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
  delenit augue duis dolore te feugait nulla facilisi.
keyword:
  - Radiosity
  - global illumination
  - constant time
firstsection: Introduction
bibliography: paper.bib
---

This template is for papers of VGTC-sponsored conferences such as IEEE
VIS, IEEE VR, and ISMAR which are published as special issues of
TVCG. The template does not contain the respective dates of the
conference/journal issue, these will be entered by IEEE as part of the
publication production process. Therefore, **please leave the
copyright statement at the bottom-left of this first page untouched**.

# Using the Style Template

(This information is incorrect for the [quarto](https://quarto.org)
template. We provide it below mostly as a markdown illustration.)

* If you receive compilation errors along the lines of `Package ifpdf
  Error: Name clash, \ifpdf is already defined` then please add a new
  line `\let\ifpdf\relax` right after the
  `\documentclass[journal]{vgtc}` call. Note that your error is due to
  packages you use that define "`\ifpdf`" which is obsolete (the
  result is that `\ifpdf` is defined twice); these packages should be
  changed to use ifpdf package instead.

* Note that each author's affiliations have to be provided in the
  author footer on the bottom-left corner of the first page. It is
  permitted to merge two or more people from the same institution as
  long as they are shown in the same order as in the overall author
  sequence on the top of the first page. For example, if authors A, B,
  C, and D are from institutions 1, 2, 1, and 2, respectively, then it
  is ok to use 2 bullets as follows:

  * A and C are with Institution 1. E-mail: {a, c}\@i1.com .
  * B and D are with Institution 2. E-mail: {b, d}\@i2.org .

* The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the "`\autoref{reference}`" call (instead of "`Figure\~{}\ref{reference}`" or similar) since "`\autoref{reference}`" turns the entire reference into an internal link, not just the number. Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.

* The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use "`\includegraphics{CypressView}`" (instead of "`\includegraphics{pictures/CypressView.jpg}`").

* For adding hyperlinks and DOIs to the list of references, you can use "`\bibliographystyle{abbrv-doi-hyperref-narrow}`" (instead of "`\bibliographystyle{abbrv}`"). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the "`http://dx.doi.org/`" url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]

  **Note 1**: occasionally (for some LaTeX distributions) this hyper-linked bibTeX style may lead to **compilation errors** ("`pdfendlink ended up in different nesting level ...`") if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to "`\bibliographystyle{abbrv-doi-narrow}`" (at the expense of removing hyperlinks from the bibliography) and try "`\bibliographystyle{abbrv-doi-hyperref-narrow}`" again after some more editing.

  **Note 2**: the "`-narrow`" versions of the bibliography style use the font "PTSansNarrow-TLF" for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the ["paratype" package](https://www.ctan.org/pkg/paratype), and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a "`-narrow`" bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the "`-narrow`" in the file name.

  DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the `abbrv-doi` bibliography style but the compiled PDF shows you the effect of the `abbrv-doi-hyperref-narrow` style.

# Bibliography Instructions

TBF.

# Example Section

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam
voluptua [@Kitware:2003,@Max:1995:OMF]. At vero eos et accusam et
justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est.

# Exposition

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat @Kindlmann:1999:SAG.

$$\sum_{j=1}^{z} j = \frac{z(z+1)}{2}$$ {#eq-1}

Lorem ipsum dolor sit amet (@eq-1), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

## Lorem Ipsum

Lorem ipsum dolor sit amet (see @tbl-vis-papers), consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum.

TBF: Finish this table.

| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: VIS/VisWeek accepted/presented papers from 1990--2016. {#tbl-vis-papers}

## Mezcal Head

Lorem ipsum dolor sit amet (see @fig-sample), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

### Duis Autem

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
ipsum dolor sit amet.

![A visualization of the 1990--2015 data from @tbl-vis-papers. The image is from @Isenberg:2017:VMC and is in the public domain.](paper-count-w-2015-new.pdf){#fig-sample width=3in}

### Ejector Seat Reservation

Duis autem by @Lorensen:1987:MCA[^1] vel eum iriure dolor in hendrerit
in vulputate velit esse molestie consequat,[^2] vel illum dolore eu
feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
ut laoreet dolore magna aliquam erat volutpat.

[^1]: The algorithm behind Marching Cubes by Lorensen et al. @Lorensen:1987:MCA had
    already been described by Wyvill et al. a
    year earlier [@Wyvill:1986:DSS].

[^2]: Footnotes appear at the bottom of the column.

#### Confirmed Ejector Seat Reservation

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat [@Nielson:1991:TAD]. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

#### Rejected Ejector Seat Reservation

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie

## Vestibulum

Vestibulum ut est libero. Suspendisse non libero id massa congue egestas nec at ligula. Donec nibh lorem, ornare et odio eu, cursus accumsan felis. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec dapibus quam vel eros mattis, id ornare dolor convallis. Donec at nisl sapien. Integer fringilla laoreet tempor. Fusce accumsan ante vel augue euismod, sit amet maximus turpis mattis. Nam accumsan vestibulum rhoncus. Aenean quis pellentesque augue. Suspendisse sed augue et velit consequat bibendum id nec est. Quisque erat purus, ullamcorper ut ex vel, dapibus dignissim erat.

Quisque sit amet orci quam. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam pharetra, nunc non efficitur convallis, tellus purus iaculis lorem, nec ultricies dolor ligula in metus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Aenean eu ex nulla. Morbi id ex interdum, scelerisque lorem nec, iaculis dui. Phasellus ultricies nunc vitae quam fringilla laoreet. Quisque sed dolor at sem vestibulum fringilla nec ac augue. Ut consequat, velit ac mattis ornare, eros arcu pellentesque erat, non ultricies libero metus nec mi. Sed eget elit sed quam malesuada viverra. Quisque ullamcorper, felis ut convallis fermentum, purus ligula varius ligula, sit amet tempor neque dui non neque. Donec vulputate ultricies tortor in mollis.

Integer sit amet dolor sit amet turpis ullamcorper varius. Cras volutpat bibendum scelerisque. Maecenas mauris dolor, gravida eu elit et, sodales consequat tortor. Integer id commodo elit. Pellentesque sollicitudin ex non nulla molestie eleifend. Mauris sagittis metus nec turpis imperdiet, vel ullamcorper nibh tincidunt. Sed semper tempus ex, ut aliquet erat hendrerit id. Maecenas sit amet dolor sollicitudin, luctus nunc sit amet, malesuada justo.

Mauris ut sapien non ipsum imperdiet sodales sit amet ac diam. Nulla vel convallis est. Etiam dapibus augue urna. Aenean enim leo, fermentum quis pulvinar at, ultrices quis enim. Sed placerat porta libero et feugiat. Phasellus ullamcorper, felis id porta sollicitudin, dolor dui venenatis augue, vel fringilla risus massa non risus. Maecenas ut nulla vitae ligula pharetra feugiat non eu ante. Donec quis neque quis lorem cursus pretium ac vulputate quam. Cras viverra tellus vitae sapien pretium laoreet. Pellentesque fringilla odio venenatis ex viverra, quis eleifend tortor ornare. Ut ut enim nunc. Vivamus id ligula nec est dignissim eleifend.

Nunc ac velit tellus. Donec et venenatis mauris. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Ut vitae lectus vel ante mollis congue. Vestibulum at cursus velit. Curabitur in facilisis enim. Vestibulum eget dui aliquet risus laoreet laoreet. Phasellus et est id magna interdum venenatis. Donec luctus vehicula justo sed laoreet. Quisque tincidunt suscipit augue, in molestie sem accumsan sed.

# Quarto Features

## Citation styles

Defend a claim by referencing a paper [@Isenberg:2017:VMC].

As Kindlmann has shown, it is possible to generate transfer functions semi-automatically [@Kindlmann:1999:SAG].

## Computational Output

```{r}
#| echo: false

# this won't be printed to the paper at all, since it produces no
# output.
library(ggplot2)
library(palmerpenguins)
```

```{r}
#| fig-pos: 't'
#| fig-width: 3
#| fig-height: 2
#| echo: false
#| warning: false
#| label: fig-penguins
#| fig-cap: Some penguins, ggplotted.

# the source code won't be printed to the paper, but the output will.
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) + geom_point()
```

<!--- this also works for Python --->
```{python}
import math
print(math.sqrt(2))
```

The penguins in @fig-penguins sure are cute.

# Conclusion

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

# Acknowledgments {.unnumbered}

The authors wish to thank A, B, and C. This work was supported in part by
a grant from XYZ (# 12345-67890).

# References {.unnumbered}